### 2025-09-23 paper/sec/related-work.tex
updated rel work

\paragraph{NLP and scientific writing.} The intersection of NLP and scientific writing has gained increasing attention in recent years, with a focus on improving the clarity, coherence, and overall quality of scientific texts. Various approaches have been proposed, including the use of automated writing assistants, text simplification techniques, and tools for enhancing the argumentative structure of scientific papers~\citep{..,}. These efforts aim to support researchers in effectively communicating their findings and engaging with the broader scientific community.

>>>>

\paragraph{NLP and scientific writing.} 
The intersection of NLP and scientific writing has gained increasing attention in the last decade, with a focus on improving the clarity, coherence, and overall quality of scientific texts. 
On the authoring side, tools such as automated writing assistants can support researchers in producing more fluent and accessible text, for instance through grammar correction, summarisation and text simplification~\citep{napoles2017jfleg,stiennon2020learning,takeshita2024cross,saggion2017automatic}. Other approaches specifically target the argumentative structure of scientific papers, helping writers to organise contributions and claims more effectively~\citep{lauscher2018arguminsci}. Nowadays, general purpose LLMs such as ChatGPT or tools tailored for the task such as Grammarly show the variety of support that NLP tools can provide to authors~\citep{wu2023chatgpt,ahn2024transformative,khalifa2024using}.

At the same time, NLP methods are being developed to assist reviewers and editors in evaluating submissions. These include systems for detecting potential issues such as lack of clarity, weak argumentative support, or even factual inconsistencies and up to scientific fraud~\citep{thakkar2025can, fromm2021argument, freedman2024detecting}. Such tools can also facilitate meta-reviewing by providing summaries of peer reviews and identifying points of disagreement among reviewers~\citep{kumar2023reviewers}. While AI tools show promises in improving the peer-review process~\citep{tyser2024ai}, there are also various risks associated such as breaches of confidentiality, lack of transparency and biases~\citep{perlis2025artificial}. Our current work situates itself in the context of NLP tools for supporting the understanding of scientific writing; specifically, it addresses one of the major critiques toward the automation of such process by offering a transparent way of examining its workflow.
